---
title: "Employer Consolidation and its Effects on Wages"
subtitle: "Evidence from hospitals shows that employer merger has effects on wages only under two certain circumstances."
author: 
  - Zhiyue Yu
  - Dingding Wang
  - Xueru Ma
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: “"
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

DOI: 10.1257 / jep.36.1.151

```{r setup, include=FALSE}
install.packages("usmap")
library(tidyverse)
```
























figure 1

```{r}
#read in data for figure1
data <- read.csv("fig_1.csv")
data
#plot figure1
figure1 <- data %>% 
  #apply ggplot
  ggplot(aes(x = year, y = brate_all)) + 
  #change the y axis name
  ylab("Births per 1,000 women age 15–44") +
  #change the line color to blue
  geom_line(color = "blue")+
  #add a vertical dot line to the graph and change the size of the line
  geom_vline(xintercept = 2007, linetype = "dotted", size = 0.8) +
  #apply minimal theme
  theme_minimal() +
  #remove the label of x-axis
  theme(axis.title.x = element_blank()) +
  geom_text(
    label="2007", 
    x=2009,
    y=69)
figure1

```








figure3
```{r}
library(haven)
data_fig3 <- read_dta("~/Downloads/package for paper 2/data/nchs/nchs_births_pop_1990_2019.dta")

#Birth Data: NCHS Microdata
temp0 <- data_fig3 %>% 
  filter((data_fig3$year >= 2004 & data_fig3$year <= 2008) | (data_fig3$year >= 2015 & data_fig3$year <= 2019)) %>%
  group_by(year, stname) %>%
  select(year, stname, numbirth1544)
  
#Pop Data: CDC SEER
age_race_comp_seer <- read_dta("~/Downloads/package for paper 2/data/pop/age_race_comp_seer.dta")
age_race_comp_seer <- age_race_comp_seer %>%
  mutate(pop1544 = pop1519 + pop2034 + pop3544) %>%
  filter( year >= 2004 & year <= 2008 | year >= 2015 & year <= 2019) %>%
    group_by(stname) %>%
  select(year, stname, pop1544)

#left join
temp0 <- left_join(temp0,age_race_comp_seer,by=c("year" = "year", "stname" = "stname"))
  
#generate new column
data_fig3 <- temp0 %>% mutate(year = case_when(year >= 2004 & year <= 2008 ~ "2004",
                                       year >= 2015 & year <= 2019 ~ "2019")) %>%
  group_by(stname,year) 

aggregated_data <- 
  aggregate(cbind(numbirth1544, pop1544)~ year + stname, data = data_fig3, FUN = mean) %>%
  mutate(brate1544_thsnds = numbirth1544/pop1544 * 1000)

library(dplyr)
#table1
table0 <- aggregated_data
table1 = table0 %>% filter(table0$year == "2004") %>% 
  rename(brate2004 = brate1544_thsnds) %>%
  select(stname, brate2004)

#table2
table2 <- aggregated_data
table3 = table2 %>% filter(table2$year == "2019") %>%
  rename(brate2019 = brate1544_thsnds) %>%
  select(stname, brate2019)

#new table
table = merge(x = table1, y = table3, by = c("stname", "stname"))
table <- table %>% 
  mutate(brate_diff = brate2019 - brate2004) %>%

  rename(state = stname)

#map
library(usmap)
library(ggplot2)
plot_usmap(data = table, values = "brate_diff") +

  scale_fill_continuous(high = "white", low = "#8B0000", name = "Population (2015)", label = scales::comma) +
   theme(legend.position = "right")


```


# Introduction


******You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).



# Data

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it. 



# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful. 

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

